Psychophysics of musical elements in the discrete-time representation of sound

### 2013-10-02 article/article.tex
ateh fonte em movimento

\subsection{Musical usage}\label{subsec:mus2}

At this point, the musical possibilities has exploded. All characteristics, like pitch (given by frequency), timbre (given by the waveform, filters and noise), volume (manipulated by intensity) and duration (given by the number of samples), can be considered in an absolute form or treated during duration of sound (with the exception of the duration itself, clearly).

@@

Another interesting possibility is to use the relations between parameters of
tremolo and vibrato, and some parameters of the basic note like frequency. In
this way, it is possible to say that vibrato frequency is proportionally
different from pitch, or the tremolo depth is inversely proportional to
pitch. Therefore, with equations \ref{vbrGamma}, \ref{vbrF} and \ref{trA}, it is
possible to define:

@@

\noindent with $f^{vbr}$ and $f^{tr}$ as $f'$ in the referenced equations. They can also be associated with the vibrato and tremolo oscillation frequency in
equation~\ref{vbrGamma}.

Besides that, $\nu$ and $V_{dB}$ are the respective depth values of vibrato and
tremolo. Functions $func_a$, $func_b$ and $func_c$ are arbitrary and dependents
on the musical intentions. The music piece \emph{Tremolos, vibratos e a
frequência} explores such characteristics and shows variations in the oscillation
waveform with specific relations with the purpose of building a \emph{musical
language}\footnote{Details in the next section.}. The corresponding code is on
Appendix~\ref{ap:tremolos} and it is also available online as part of the \massa
toolbox.

About convolution it is possible to use a musical pulse as duration -- like a BPM
pulse -- and to distribute impulses during this original pulse, aiming at
establishing metrics and rhythms\footnote{It is important to remember that the
convolution with a impulse results in a sound shifted to the moment that the
impulse occurred}.
For example, two impulses equally spaced builds a binary division into the
pulse. Two signals, one with 2 pulses and another with 3 pulses, both with
equally spaced impulses in the pulse duration, results in the pulse
maintenance with a rhythmic mark that is possible to use as binary or ternary
divisions in many ethnic or traditions music styles~\cite{Gramani}. 
The absolute values of the impulses results in proportions among the
amplitudes of the convoluted signals.
Using convolution with impulses as a metric is explored in the
music piece \emph{Trenzinho de caipiras impulsivos}. The features
embrace the creation of 'sound amalgam' based on granular synthesis and this
piece provides a link to the next section. Refer specially to the
Figure~\ref{fig:pulsoSubAgl}. The source code of the music piece is included in the section~\ref{ap:trenzinho} and online as well (\massa toolbox).

>>>>

\subsection{Musical usages}\label{subsec:mus2}

At this point, the musical possibilities exploded. All characteristics, like pitch (given by frequency), timbre (given by the waveform, filters and noise), volume (manipulated by intensity) and duration (given by the number of samples), can be considered in an absolute form or treated during duration of sound (clearly, with the exception of duration itself).

@@

An interesting possibility is to use relations between parameters of
a tremolo or vibrato, and some parameters of the basic note like frequency. In
this way, it is possible to have a vibrato frequency proportional
to note pitch, or a tremolo depth inversely proportional to
pitch. Therefore, with equations \ref{vbrGamma}, \ref{vbrF} and \ref{trA}, it is
perfectly possible to arbitrate:

@@

\noindent with $f^{vbr}$ and $f^{tr}$ as $f'$ in the referenced equations. They can also be associated with vibrato and tremolo oscillation frequency of
equation~\ref{vbrGamma}. 
 $\nu$ and $V_{dB}$ are the respective depth values of vibrato and
tremolo. Functions $func_a$, $func_b$ and $func_c$ are arbitrary and dependent
on musical intentions. The music piece \emph{Tremolos, vibratos e a
frequência} explores such bonds and exhibit variations in the oscillation
waveform with the purpose of building a \emph{musical
language} (details in the next section). The corresponding code is on
is also available online as part of the \massa\ toolbox.

\subsubsection{Convolution for rhythm and meter}
It is possible to use a musical pulse -- such as specified by a BPM
tempo mark -- and to place an impulse at the start of each beat, with the purpose of
establishing metric and rhythms: the
convolution with an impulse shifts the sound to where
the impulse.
For example, two impulses equally spaced builds a binary division into the
pulse. Two signals, one with 2 impulses and another with 3 impulses, both 
equally spaced in the pulse duration, yields the pulse
maintenance with a rhythmic trace in which is easy to use binary or ternary
divisions of the pulse, as found in many ethnic or traditional music styles~\cite{Gramani}. 
Absolute values of the impulses implyes in proportions among the
amplitudes of the sonic incidences.
The use of convolution with impulses in this context is explored in the
music piece \emph{Trenzinho de caipiras impulsivos}. These features
embrace the creation of `sound amalgams' based on granular synthesis, for now refer specially to the
figure~\ref{fig:pulsoSubAgl}. This piece is a link to next section contents. 
The source code of the music piece is online, as part of the \massa toolbox.
